---
title       : Storm Watch
subtitle    : Visual Storm Display
author      : 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, quiz, bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Storm Watch 

* A Storm Visualizer for the United States
* Allows you to visualize the distribution of Inclement Weather across the country
* Broken Down by State, Event Type, Month, Fatality Count
* Helps you and your organization company plan events

--- .class #id 

## Event Occurances By Month

* Storm Watch allows you to display the breakdown of events by Month
* Enables planning of events by the calendar

```{r,echo=FALSE,message=FALSE,cache=TRUE}
library(shiny)
library(lubridate)

data(state)
old_storms <- read.csv("result2010.txt",strip.white=TRUE,fill=TRUE,nrow=-1)
col_names <- c("BGN_DATE","STATE","EVTYPE","FATALITIES","LATITUDE","LONGITUDE" )
colnames(old_storms) <- col_names
old_storms$Date <- as.Date(sapply(strsplit(as.character(old_storms$BGN_DATE),' '),
                              function(x) {x[[1]]}),format = '%m/%d/%Y')

old_storms$Year <- year(old_storms$Date)
old_storms$Month <- months.Date(old_storms$Date)
old_storms$MonthF <- factor(old_storms$Month,levels=month.name)
#old_storms$MonthF <- factor(old_storms$Month,levels=(month.abb(old_storms$Month)))
#print("Date Complete")
old_storms$EVTYPE <- toupper(old_storms$EVTYPE)
old_storms$EVTYPEN <- "OTHER"
old_storms$EVTYPEN[grep("*TORNADO*", old_storms$EVTYPE)] <- "TORNADO"
old_storms$EVTYPEN[grep("*HURRICANE*|*TYPHOON*", old_storms$EVTYPE)] <- "HURRICANE"
old_storms$EVTYPEN[grep("*WIND*", old_storms$EVTYPE)] <- "WIND"
old_storms$EVTYPEN[grep("*FIRE*", old_storms$EVTYPE)] <- "FIRE"
old_storms$EVTYPEN[grep("*STORM*|*GLAZE*|*HAIL*|*WETNESS*|*LIGHTNING*|*RAIN*|*BLIZZARD*", old_storms$EVTYPE)] <- "STORM"
#old_storms$EVTYPEN[grep("*COLD*|*LOW TEMPERATURE*|*WINTRY*|*FREEZE*", old_storms$EVTYPE)] <- "COLD"
#old_storms$EVTYPEN[grep("*SNOW*", old_storms$EVTYPE)] <- "SNOW"
old_storms$EVTYPEN[grep("*FLOOD*|*STREAM*|*HIGH WATER*", old_storms$EVTYPE)] <- "FLOOD"
old_storms$EVTYPEN[grep("*HEAT*|*HOT*", old_storms$EVTYPE)] <- "HEAT"
old_storms$EVTYPEN[grep("*SURF*|*SEAS*|*MARINE*|*CURRENT*|*TSUNAMI", old_storms$EVTYPE)] <- "SURF"
old_storms$EVTYPEN[grep("*FOG*", old_storms$EVTYPE)] <- "FOG"
old_storms$EVTYPEN[grep("*DRY*|*DROUGHT*", old_storms$EVTYPE)] <- "DRY"
old_storms$EVTYPEN[grep("*LANDSLIDE*|*LAND*|*AVALANCHE*|*SLIDE*", old_storms$EVTYPE)] <- "LANDSLIDE"
old_storms$EVTYPEN[grep("*ICE*|*ICY*|*FROST*|SNOW*|*COLD*|*LOW TEMPERATURE*|*WINTRY*|*FREEZE*", old_storms$EVTYPE)] <- "COLD"

#storms <- na.omit(old_storms)
old_storms$lat <- as.numeric(as.character(old_storms$LATITUDE))/100.0
old_storms$long <- (as.numeric(old_storms$LONGITUDE)/100.0)*-1.0
storms <- old_storms[as.character(old_storms$STATE) %in% state.abb,]
storms <- storms[complete.cases(storms),]
storms$REGION <- state.region[match(storms$STATE,state.abb)]
rm(old_storms)

valid_events  <- c("TORNADO","FLOOD","STORM")        
storms$EVTYPESTRING = paste(storms$EVTYPEN,storms$Date,"-")
storms <- storms[as.character(storms$EVTYPEN) %in% valid_events,]


state_list <- as.list(sort(as.character(unique(storms$STATE))))

library(ggplot2)
p1 <- ggplot(data=storms,aes(x=MonthF,fill=EVTYPEN)) + geom_bar() + scale_size_area() + 
  xlab("Month") +
  ylab("Events") 
p1

```

--- .class #id 

## Event Occurances by State

* Storm Watch allows you to display the breakdown of events by State
* Enables facilities planning based on weather risks involved

```{r,echo=FALSE,cache=TRUE}
p2 <- ggplot(data=storms,aes(x=EVTYPEN,fill=STATE)) + geom_bar() + scale_size_area() + 
  xlab("Event Type") +
  ylab("States") 
p2

#p2 <- ggplot(data=storms,aes(x=EVTYPEN,fill=STATE)) + geom_bar() 
#p2
```

--- .class #id 

## Table of Event Occurances 

* A regional summary of events
* Because nature of inclement weather is regional

```{r, echo=FALSE,cache=TRUE,message=FALSE}
library(datasets)
library(plyr)
library(dplyr)  
library(tidyr)


# Run the functions length, mean, and sd on the value of "change" for each group, 
# broken down by sex + condition
require(dplyr)
stormsG <- storms %>%
  group_by(EVTYPEN,REGION) %>%
  summarise(Counts = n()) %>%
  tidyr::spread(EVTYPEN,Counts)

stormsG

```

--- .class #id 

## Table of Event Occurances 

* A monthly calendar broken down by region

```{r, echo=FALSE,cache=TRUE,message=FALSE}
library(datasets)
library(plyr)
library(dplyr)  
library(tidyr)


require(dplyr)
stormsM <- storms %>%
  group_by(REGION,MonthF) %>%
  summarise(Counts = n()) %>%
  tidyr::spread(REGION,Counts)

stormsM <- stormsM[order(stormsM$MonthF), ]
stormsM

```

--- .class #id 
